### 2023-02-03 outputs/paper/Paper 1.qmd
Toronto Victimization rate: second commit

---
title: "Assessing the Rates of Crimes Against the Person in Toronto "
author: Sofia Sellitto
date: February 3, 2023
date-format: long
abstract: Abstract 
format: pdf
bibliography: bibliography.bib
---

#

This data was free for use, courtesy of @opendatatoronto. We are using @R for analyzing this data.

## Loading Packages and Data

We import tidyverse by @tidyverse for convenience because it contains many other packages which will be useful for plotting and data manipulations. We also use @janitor for cleaning and @dplyr for manipulating data.

```{r}
## loading appropriate packages

library(tidyverse)
library(ggplot2)
library(knitr)
library(janitor)
library(dplyr)

```

```{r}
## Calling on cleaned data 
cleaned_data_dropped <- read_csv(here::here("inputs/data/cleaned_data.csv"))
head(cleaned_data_dropped)
```

```{r}

## summing count of crimes and making them its own group

cleaned_data_dropped_grouped <- cleaned_data_dropped |>
  group_by(subtype, reported_year) |>
  summarise(across(count, sum))

cleaned_data_dropped_grouped
```

```{r}
### making a data frame that includes the variable sex so that I can call on it for a graph

sex_data <- cleaned_data_dropped |>
  group_by(subtype,sex, reported_year) |>
  summarise(across(count, sum))

sex_data

```

```{r}

## calculating the mean number of crimes for each subcategory across 2014-2021

mean_data <- cleaned_data_dropped_grouped |>
    summarise(mean_count = mean(count))

mean_data

```

```{r}

## Creating a table to visually to display the mean number of crimes for each subcategory across 2014-2021

knitr::kable(mean_data,"simple", caption = "Average Number of victims of crimes in Toronto across 2014-2021, grouped by category of crime", 
             col.names=c("Category of Crime", "Average Number of Victims"), 
             align=c('l', 'c'))

```

```{r}

knitr::kable(head(cleaned_data_dropped_grouped), "simple", caption = "The first 6 rows of the data frame used to represent the number of victims according to category of crime in Toronto Across 2014 - 2021",
             col.names=c("Sub-Category of Crime", "Year", "Number of Victims"), 
             align=c('l', 'c', 'c'))

```

```{r}

knitr::kable(head(sex_data), "simple", caption = "The first 6 rows of the data frame used to represent the gender of the number of victims according to category of crime in Toronto Across 2014 - 2021", col.names=c("Sub-Category of Crime", "Sex", "Year", "Number of Victims"), 
             align=c('l', 'c', 'c'))

```

See @figure1

```{r}
## creating a line graph that includes the number of victims according to category of crime in Toronto across 2014-2021

#| fig-cap: "Number of victims according to category of crime in Toronto Across 2014 - 2021"
#| label: figure1

cleaned_data_dropped_grouped |>
  ggplot(aes(x = reported_year, y = count, colour = subtype))+
  geom_point(linewidth = 5, alpha = 0.3)+
  geom_line(linewidth = 1)+
  theme_minimal()+
  labs(
    title = "Victims of Crime in Toronto Across 2014 - 2021",
    x = "Year",  
    y = "Number of Victims", 
    colour = "Category of Crime" )

```

See @figure2

```{r}

## calling on my data frame to create a bar plot of the sex of the victims across time

#| fig-cap: "How the victims of crimes varries across sex in Toronto from 2014-2021"
#| label: figure2

ggplot(sex_data, aes
       (x =reported_year, 
        y =count, 
        fill =sex))+
            geom_bar(position="dodge", stat="identity")+
            facet_wrap(~subtype)+
            scale_fill_manual(values=c("#C0AFE2", "#ADD8E6"))+
            theme_minimal()+
            labs(
            title = "Victims of Crime in Toronto Categorized by Sex Across 2014 - 2021",
            x = "Year",  
            y = "Number of Victims",
            legend = "Sex")

```

## References

### 2023-02-03 inputs/data/00-download_data.R

@@ -1,0 +1,19 @@
+
+library(opendatatoronto)
+library(dplyr)
+
+# get package
+package <- show_package("police-annual-statistical-report-victims-of-crime")
+package
+
+# get all resources for this package
+resources <- list_package_resources("police-annual-statistical-report-victims-of-crime")
+
+# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+
+# load the first datastore resource as a sample
+data <- filter(datastore_resources, row_number()==1) %>% get_resource()
+data
+
+write.csv(data, "~/Inputs/Tutorial 1/Input/raw_data.csv", row.names=FALSE)

### 2023-02-03 scripts/00-cleaning.R

@@ -1,0 +1,59 @@
+data <- read.csv( "~/Inputs/Tutorial 1/Scripts/raw_data.csv")
+
+## checking for all possible categories of crime included in the data set 
+
+unique(data$Subtype)
+
+
+## checking for all possible genders included in the data set 
+
+unique(data$Sex)
+
+## making names easier to type  
+
+cleaned_data <-
+  clean_names(data)
+
+head(cleaned_data)
+
+
+
+## Checking for missing data in year column, if a number is given back, we know no dates are missing  
+
+mean(cleaned_data$reported_year)
+
+
+## Cross-tabulating to understand the pattern of unknown data
+
+xtabs(~sex+age_cohort, data=cleaned_data)
+
+
+## Changing "count" column from a character to numeric so it can be summed
+
+cleaned_data_dropped <- cleaned_data 
+
+
+cleaned_data_dropped$count = as.numeric(as.character(cleaned_data_dropped$count))
+
+head(cleaned_data_dropped)
+
+## Dropping 'U' (unknown) from the sex variable
+cleaned_data_dropped <- cleaned_data_dropped[!(cleaned_data_dropped$sex=='U'),]
+
+## Selecting only the variables of interest 
+
+cleaned_data_dropped <- cleaned_data_dropped |> 
+  select(
+    reported_year,
+    subtype,
+    sex,
+    count
+    
+  )
+
+
+
+
+write.csv(cleaned_data_dropped, "~/Inputs/Tutorial 1/Input/cleaned_data.csv", row.names=FALSE)
+
+

### 2023-02-03 scripts/00-simulation.R

@@ -1,0 +1,30 @@
+## Simulating how I wish to display my data 
+
+years <- sample(x = 2014:2021,size = 4,replace = TRUE)
+
+crimes <- c("Assault", "Robbery", "Sexual Violence", "Other")
+    
+sexs <-  c("M", "F", "F", "M")
+    
+number_victims <- sample(x = 1:1700,size = 4,replace = TRUE)
+
+df_s = data.frame(
+ Year = years,
+ Crimes = crimes,
+ Sex = sexs,
+ Victims = number_victims
+)
+
+#Test 
+
+max(df_s$number_victims)
+min(df_s$number_victims)
+
+any(df_s$number_victim%%1!=0)
+
+max(df_s$years)
+min(df_s$years)
+
+any(df_s$years%%1!=0)
+
+

### 2023-02-03 outputs/paper/Paper 1.qmd
Toronto Victimization: third commit with a more organized and complete structure

@@
 ---
-title: "Assessing the Rates of Crimes Against the Person in Toronto "
+title: " Sin and The City: Investigating the Number and Sex of Victims of Crimes Against the Person in Toronto Across 2014-2021. "
 author: Sofia Sellitto
@@
 date-format: long
-abstract: Abstract 
-format: pdf
+format: 
+  pdf: 
+    documentclass: article 
+    geometry: margin = 1in
+    abstract: In the city of Toronto, citizens have become accustomed to fearing for their safety as crime appears to have increased in recent years.
+In this paper, I analyze the number of victims of crimes against the person in the City of Toronto from 2014 - 2021 to investigate if the number of Torontonians victimized by crime has increased over time and if sex differences are apparent across various categories of crime.
+By examining the data I found that in all categories of crime (assault, robbery and other) but sexual assault, the number of annual victims has decreased over time, sexual violence appears to have increased in 2021 maintaining the largest sex difference, skewing drastically female.
+These results have implications for how Torontonians view their safety and consider how their sex impacts the possibility of crimes they fall victim to.
+toc: true
+toc-depth: 4
+theme: cosmo
 bibliography: bibliography.bib
@@
 
-#
+Code and data supporting this analysis is available at: LINK
 
-This data was free for use, courtesy of @opendatatoronto. We are using @R for analyzing this data.
+## 1 Introduction
 
-## Loading Packages and Data
+Prior to the 2006 Canadian federal election, Stephen Harper, who would eventually be elected Prime Minister until 2016, led a conservative campaign based on crime prevention (cite SH book).
+The conservative party asserted that Canada was facing a period of increased crime and assured protection to Canadians by way of tougher sentencing/ parole, additional police funding, and investments in victim services, among a myriad of other things [@book].
+In his 2011 campaign, Harper suggested that judicial opinions favoured the rights of criminals instead of Canadian citizens [@book].
+Harper's claims of increased crime and strides for harsher punishment sparked debates of fear-mongering and inducing an era of incorrectly generated moral panic.
+According to statistics Canada, the total rate of crime had declined over 20 years since its peak in 1991 [@book].
+This was in direct contradiction to Harper's claims.
 
-We import tidyverse by @tidyverse for convenience because it contains many other packages which will be useful for plotting and data manipulations.
-We also use @janitor for cleaning and @dplyr for manipulating data.
+Less than a decade following the Harper era, in the city of Toronto, violent crimes against the person have become customary as Torontonians regularly hear about and fear crime.
+Many of the crimes that have come to shake Torontonians, specifically those which take place on the major transportation organization, the TTC (Toronto Transit Commission) appear to target random citizens and are often extremely violent forms of assault.
+Akin to Harper, it is challenging not to believe that crime has increased and feel at risk of being a victim.
+Nevertheless, as a young woman living in Toronto's Downtown core, it is difficult to avoid nihilistic thinking and continual fears concerning my safety as I exist in this city.
+While the analysis of cumulative crime rates is extremely valuable, I don't believe that this represents the root of Torontonian's fear - the possibility of being a victim themselves.
+
+In this paper, I utilize Open Data Toronto [@opendatatoronto] to assess the number of victims of crimes against the person in Toronto across 2014 - 2021.
+This is to investigate how the number of victims of crime in Toronto has changed over time, and if it has increased  over time, which categories of crimes against the person are most prevalent and if sex differences are apparent across different forms of crime.
+I will commence with a discussion of my data source and its features.
+I will then conduct a breakdown of the average number of victimization reports for the sub-categories "assault", "robbery", "sexual violence" and "other", across 2014-2019.
+Next, I will move towards an analysis of how the reported number of victims has changed according to the category of crime, over time.
+I will conclude my analysis with an investigation of how the number of victimization reports for each category of crime, differs according to the gender of the victim.
+The paper will culminate with a discussion of my findings, limitations and future research possibilities.
 
 ```{r}
+#| echo: false
+#| warning: false
+
 ## loading appropriate packages
@@
 ```{r}
+#| echo: false
+#| warning: false
+#| include: false
+
 ## Calling on cleaned data 
@@
 ```{r}
+#| echo: false
+#| warning: false
+#| include: false
 
@@
 ```{r}
+#| echo: false
+#| warning: false
+#| include: false
+
 ### making a data frame that includes the variable sex so that I can call on it for a graph
@@
 
+## 2 Data
+
+#### 2.1 Data Source
+
+The data source utilized for the paper was taken from the City of Toronto's open data portal [@opendatatoronto].
+The data was collected from the Toronto Police Service Annual Statistical Report (ASR) which conducts overviews of police-related statistics.
+For this paper specifically, the dataset is comprised of all identified victims of crimes against the person (victims of crimes) from 2014-2021.
+
+#### 2.2 Data Features
+
+The data was collected across 8 years, spanning from 2014-2021.
+It includes all reports of identified victims of crimes against the person found within but not limited to the city of Toronto (some crimes may have occurred outside of city limits) and are not verified by location.
+
+\
+The data is categorized across 9 variables which include; the reported year of the crime, the category, category subtype and assault subtype of the crime, the sex of the victim(s), their age group and cohort, and the count (number) of identified victims.
+It was stated that duplicates concerning the count of identified victims are possible if the same individual was victimized for the same crime during the same period.
+This leads me to a discussion of potential biases included in the data source.
+To begin, it can often be difficult to fully trust that police statistics were ethically collected and reported.
+Oftentimes, citizens make statements that police fail to report or later on dismiss.
+Furthermore, the Toronto police may hold biases against communities of colour, thus potentially making them more likely to create victim reports against wrongfully accused individuals of colour.
+
+\
+This paper will follow an analysis of the following variables: the reported year of the crime, the crime category subtype, the number of identified victims and their respective sex.
+I have chosen to exclude the assault subtype of the crime as it included numerous missing values.
+I would have been forced to remove this data, thus negatively impacting my analysis.
+Because my analysis is focused on victims of crime over time and potential gender differences, I have decided to withdraw from using data for age group or cohort.
+
+\
+Included within the crime category subtype are (1) assault, (2) robbery, (2) sexual violation and (4) other.
+To further discuss these categorizations, in Canada, assault is classified as intentionally applying force to an individual indirectly or directly without consent.
+This can then include assault to cause bodily harm, with a deadly weapon, aggravated assault or attempted murder.
+In Canada, robbery is defined as stealing for extortion of the article stolen and can include the use of violent threats, personal violence or with a weapon.
+In Ontario, sexual violence describes any physical or psychological violence conducted or targeted in a sexual means.
+Unfortunately, no information was provided by the City of Toronto nor the Toronto Police Service Annual Statistical Report (ASR) on the nature of the "other" category.
+We are forced to assume that this may include break and enter, homicide, domestic violence or various forms of crimes against the person.
+
+#### 2.3 Data Analysis
+
+Using the R coding language [@R], I processed, cleaned and analyzed the data set .
+I then loaded the packages tidyverse [@tidyverse], janitor [@janitor] and dplyr [@dplyr].
+These were used throughout my analysis to call on, sort, categorize and clean my data.
+To display my data, the packages ggplot2 [@ggplot] and knitr [@knitr] were loaded and used respectively for creating graphs and tables.
+
+I initiated my analysis by cleaning and sorting my data.
+I first selected the variables of interest, checked for missing values, cross-tabulated to understand the pattern of unidentified data and dropped unknown variables.
+Included in the sex variable was "U" for unknown, which unfortunately left me forced to remove it from my dataset.
+This is a limitation of my analysis that I will further examine in my discussion.
+
+My investigation began with an overview of the average number of people victimized by crime in Toronto across 2014-2021 based on the sub-category of crime.
+My intent with this is to help visualize the numerical average of Torontonians who were the victims of each sub-category of crime across 2014 - 2021.
+I visualized this in a table (see table 1) to create categorical organization from the highest to lowest average victims of crime.
+
 ```{r}
+#| echo: false
+#| warning: false
 
@@
 
-knitr::kable(mean_data,"simple", caption = "Average Number of victims of crimes in Toronto across 2014-2021, grouped by category of crime", 
+knitr::kable(mean_data,
+             "simple", 
+             caption = "Average Number of victims of crimes in Toronto across 2014                         -2021, grouped by category of crime", 
              col.names=c("Category of Crime", "Average Number of Victims"), 
@@
 
+From there, I began preparing the data frame I would need to visualize how the number of victims in each sub-category of crime has changed over time.
+I included the 3 variables; sub-categories, years and the number of victims.
+Because my data frame still remained relatively large, below, I have visualized the data in a table (see Table 2), which presents only the **first 6 rows** of the data frame I created to be called on later in my analysis.
+
 ```{r}
+#| echo: false
+#| warning: false
 
-knitr::kable(head(cleaned_data_dropped_grouped), "simple", caption = "The first 6 rows of the data frame used to represent the number of victims according to category of crime in Toronto Across 2014 - 2021",
+knitr::kable(head(cleaned_data_dropped_grouped), 
+             "simple", 
+             caption = "The first 6 rows of the data frame used to represent the                          number of victims according to category of crime in                                Toronto Across 2014 - 2021",
              col.names=c("Sub-Category of Crime", "Year", "Number of Victims"), 
@@
 
+Similarly, to create the data frame that I would use to visualize how sex impacts the victims of crimes in Toronto, I created a data frame, akin to the one above, but instead included the variable 'sex'.
+Again, the data frame remained rather large, so below I have visualized the data in a table (see Table 3), which presents the **first 6 rows** of the data frame I created to call on in my analysis.
+
 ```{r}
+#| echo: false
+#| warning: false
 
-knitr::kable(head(sex_data), "simple", caption = "The first 6 rows of the data frame used to represent the gender of the number of victims according to category of crime in Toronto Across 2014 - 2021", col.names=c("Sub-Category of Crime", "Sex", "Year", "Number of Victims"),
+knitr::kable(head(sex_data), 
+             "simple", 
+             caption = "The first 6 rows of the data frame used to represent the                           gender of the number of victims according to category of                           crime in Toronto Across 2014 - 2021", col.names=c("Sub                             -Category of Crime", "Sex", "Year", "Number of Victims"),
              align=c('l', 'c', 'c'))
@@
 
-See @figure1
+## 3 Results
+
+As per my primary objective for this paper, I wish to understand how the number of victims of crime in Toronto has changed over time and to analyze if victimization numbers are increasing over time.
+Because of this, I created a line graph that showcases the number of victims according to each subcategory of crime over 2014 -2019 (See Figure 2).
+To do this, I called on the data frame visualized above (See Table 2).
+This visualization allows for comparison across the different subcategories of crimes while simultaneously presenting a rich depiction of how the number of victims has changed over time, the year they peaked and at their lowest.
+Using the data previewed in Table 1, I was able to discriminate the number of total victims for the peak years of each sub-category of crime.
+
+See **Figure 1**
 
 ```{r}
+#| echo: false
+#| label: figure1
+#| fig-cap: "Number of victims according to sub-category of crime in Toronto Across 2014 - 2021"
+#| warning: false
+
 ## creating a line graph that includes the number of victims according to category of crime in Toronto across 2014-2021
-
-#| fig-cap: "Number of victims according to category of crime in Toronto Across 2014 - 2021"
-#| label: figure1
 
@@
   labs(
-    title = "Victims of Crime in Toronto Across 2014 - 2021",
+    title = "The Number of Victims of Crime in Toronto Across 
+2014 - 2021",
     x = "Year",  
     y = "Number of Victims", 
-    colour = "Category of Crime" )
+    colour = "Category of Crime" )+
+    theme(plot.title=element_text(hjust=0.5))
 
@@
 
-See @figure2
+By drawing your attention to Figure 1, it is apparent from 2014-2021, the most prevalent crime that Torontonians were victims of was "Assault".
+From 2014 - 2018, the number of annual victims gradually increased until it peaked in 2019, with 18819 total victims.
+The number of victims of assault then began to decrease following 2019, where it spiked in 2021 but remained below 17500 victims.
+
+The "Other" category held the second-highest number of victims from 2014 - 2021.
+Unfortunately, due to the nature of this variable, valuable insights cannot be drawn.
+However, it peaked in 2015, seeing 5605 victims, where it steadily decreased until 2021.
+
+"Robbery" had the third most victims of crime from 2014-2021.
+The peak number of victims for this variable is not very distinct, but saw its highest in 2017 with 3686 victims that year.
+From 2017 onward the number of victims of robberies steadily decreased.
+
+"Sexual Violence" had the least number of victims of crime from 2014-2021.
+However, contrary to the others, sexual violence saw its peak in 2021 with 2568 victims.
+From 2014 onward, the number of victims gradually increased until it peaked in 2021.
+
+In representing how sex impacts the victims of crimes, I chose to display this data using 4 bar charts based on the categories "Assault", "Other", "Robbery" and "Sexual Violence".
+Each plot assesses the number of victims across 2014 - 2021 and the number of victims that were male (M) and female (F), respectively.
+Using the data set from Table 2, which was previewed above, I determined the number of victims that were male and female for the peak years of each sub-category of crime and calculated percentage differences.
+
+\
+See **Figure 2**
 
 ```{r}
+#| echo: false
+#| label: figurew
+#| fig-cap: "How the victims of crimes varries across sex in Toronto from 2014-2021"
+#| warning: false
 
 ## calling on my data frame to create a bar plot of the sex of the victims across time
-
-#| fig-cap: "How the victims of crimes varries across sex in Toronto from 2014-2021"
-#| label: figure2
 
@@
             labs(
-            title = "Victims of Crime in Toronto Categorized by Sex Across 2014 - 2021",
+            title = "The Number of Victims of Crime in Toronto by  
+Sex Across 2014 - 2021",
             x = "Year",  
             y = "Number of Victims",
-            legend = "Sex")
+            legend = "Sex") +
+            theme(plot.title=element_text(hjust=0.5))
 
@@
 
+In drawing your attention to Figure 2 contained in the "Assault", on the top left, it is apparent that from 2014 - 2021, more male Torontonians were victims of assault than women.
+In its peak year, 2017, of the 18819 victims, 8679 were women and 10140 were men.
+This would mean that in 2017, there was 1461 more male than female victims, with a difference of 15.53%.
+
+According to "Other", on the top right, it is apparent that from 2014 - 2021, more female Torontonians were victims than men.
+Again, because of the nature of this variable, insightful analysis cannot be made, however, the difference between male and female victims is minimal.
+
+As per "Robbery", on the bottom left, it is apparent that from 2014 - 2021, more male Torontonians were victims of robbery than women.
+Of the 3686 victims in its peak year, 2017, 909 victims were female, and 2777 were men.
+This would mean that in 2017, there were 1868 more male victims than females, with a difference of 101.35%.
+
+\
+Lastly, "Sexual Violence", on the bottom right, shows the most apparent sex difference from 2014-2021, where more female Torontonians were victims of sexual violence than men.
+This sex difference was the most defined across all the sub-categories, with 2244 of the total 2568 victims in its peak year, 2021 being female, and 334 being male.
+This means a difference of 147.9%.
+
+## 4 Discussion
+
+#### 4.1 Findings
+
+In this paper, I have utilized The City of Torontos open data portal, supplied by the Toronto Police Service Annual Statistical Report (ASR) to assess the number of victims of crimes against the person in Toronto across 2014-2021.
+The primary objective of this paper was to assess if the number of Torontonians victimized by crime has changed and if it has increased over time, which categories of crimes against the person are most prevalent and if sex differences are apparent across the categories of crime.
+By utilizing the Coding language R (cite) I was able to import, clean and asses the raw data made available by the City.
+In doing so I discovered that for the sub-category of crimes, assault, robbery, and other, the number of Torontonian victims peaked in the years 2019, 2017, and 2015 respectively and decreased until the end of the data collection date (2021).
+The most prevalent crime that Torontonians were victims of was assault.
+Contrarily, sexual violence saw its peak in 2021, meaning it has not decreased over time.
+In assessing how sex affects the number of victims for each sub-category of crime, it was apparent that sexual violence had the largest difference between male and female victims, skewing drastically female, maintaining a difference of 147.97%.
+Robbery maintains the second largest difference between male and female victims, skewing towards men with a difference of 101.35%.
+"Other", had the third largest difference, with more female victims than male.
+Assault had the least large sex difference, with slightly more male victims than feel with a difference of 15.53%.
+
+#### 4.2 Limitations
+
+Because of the nature of the data set, my analysis is limited in many regards.
+To begin with the source, Torontos Police Service Annual Statistical Report (ASR), it is possible that data was misrepresented, incorrectly input, missing, duplicated or potentially falsified.
+While it is important to avoid speculation about these statistics' validity, I believe it is necessary to highlight possibilities for error.
+Another vital point to acknowledge is the number of individuals who may have fallen victim to a crime against the person and failed to report it or those who intended to file a report but were dismissed.
+The latter is an extremely apparent phenomenon for women who attempt to report sexual violence.
+While this data represents the number of individuals who have fallen victim to crimes against the person across 2014-2021, it cannot account for unfiled and dismissed reports.
+Meaning that the number of victims could potentially be far higher than portrayed in the analysis.
+Regarding the data set itself, removing missing and unknown data will impact the analysis in various regards.
+More specifically, regarding the variable sex variable including "male", "female" and "unknown", I was forced to remove "unknown" as further insights were not provided on this denomination.
+Removing this data could have prevented insightful analysis of non-binary and Agender individuals.
+
+#### 4.3 Future Research
+
+In the future, I believe that efforts should be made to analyze and investigate data on crime and victimization rates for the most updated reports from 2022-2023.
+As mentioned earlier in my paper, throughout 2022 and 2023, assaults occurred in Toronto's major transportation commission, the TTC.
+In the media, it appeared as though these crimes were novel and occurring far more frequently than ever before.
+While my results showed that the number of assault victims peaked in 2019, and began to decline until 2021, it is possible that this data looks different when accounting for the current day.
+Furthermore, I believe that future research on the rate of crime and victims on the TTC would allow for insightful and meaningful findings for Torontonians.
+
+\
+
 ## References

### 2023-02-03 scripts/00-cleaning.R

@@ -1,59 +1,49 @@
 data <- read.csv( "~/Inputs/Tutorial 1/Scripts/raw_data.csv")
 
 ## checking for all possible categories of crime included in the data set 
-
 unique(data$Subtype)
 
 
-## checking for all possible genders included in the data set 
-
+## checking for all possible sexs/ genders included in the data set 
 unique(data$Sex)
 
-## making names easier to type  
 
+## making names easier to type  
 cleaned_data <-
   clean_names(data)
-
 head(cleaned_data)
 
 
-
 ## Checking for missing data in year column, if a number is given back, we know no dates are missing  
-
 mean(cleaned_data$reported_year)
 
 
 ## Cross-tabulating to understand the pattern of unknown data
-
 xtabs(~sex+age_cohort, data=cleaned_data)
 
 
 ## Changing "count" column from a character to numeric so it can be summed
-
 cleaned_data_dropped <- cleaned_data 
-
-
 cleaned_data_dropped$count = as.numeric(as.character(cleaned_data_dropped$count))
-
 head(cleaned_data_dropped)
+
 
 ## Dropping 'U' (unknown) from the sex variable
 cleaned_data_dropped <- cleaned_data_dropped[!(cleaned_data_dropped$sex=='U'),]
 
-## Selecting only the variables of interest 
 
+## Selecting only the variables of interest 
 cleaned_data_dropped <- cleaned_data_dropped |> 
   select(
     reported_year,
     subtype,
     sex,
     count
-    
   )
 
 
 
 
-write.csv(cleaned_data_dropped, "~/Inputs/Tutorial 1/Input/cleaned_data.csv", row.names=FALSE)
+write.csv(cleaned_data_dropped,"inputs/data/cleaned_data.csv", row.names=FALSE)
 
 

### 2023-02-03 scripts/00-simulation.R

@@ -1,4 +1,6 @@
-## Simulating how I wish to display my data 
+## Simulating ##
+
+## Creating variables and their possible values/categories
 
 years <- sample(x = 2014:2021,size = 4,replace = TRUE)
 
@@ -8,6 +10,8 @@ sexs <-  c("M", "F", "F", "M")
     
 number_victims <- sample(x = 1:1700,size = 4,replace = TRUE)
 
+## Making data frame out of my simulated columns 
+
 df_s = data.frame(
  Year = years,
  Crimes = crimes,
@@ -15,16 +19,20 @@ df_s = data.frame(
  Victims = number_victims
 )
 
-#Test 
+## Testing ##
 
-max(df_s$number_victims)
-min(df_s$number_victims)
+### Checking that values are whole numbers for both "Year" and "Number of Victims"
 
 any(df_s$number_victim%%1!=0)
 
-max(df_s$years)
-min(df_s$years)
-
 any(df_s$years%%1!=0)
+
+
+### Checking for type of number of unique types
+
+unique(crimes, incomparables = FALSE)
+
+unique(sexs, incomparables = FALSE)
+
 
 

### 2023-02-03 outputs/paper/Paper 1.qmd
Toronto Victimization: Fourth commit, includes updated README, and output documents such as paper 1.qmd and paper 1.pdf

@@
 ---
-title: " Sin and The City: Investigating the Number and Sex of Victims of Crimes Against the Person in Toronto Across 2014-2021. "
+title: " Sin and The City: Investigating the Number and Sex of Victims of Crimes Against the Person in Toronto Across 2014-2021."
 author: Sofia Sellitto
@@
 In this paper, I analyze the number of victims of crimes against the person in the City of Toronto from 2014 - 2021 to investigate if the number of Torontonians victimized by crime has increased over time and if sex differences are apparent across various categories of crime.
-By examining the data I found that in all categories of crime (assault, robbery and other) but sexual assault, the number of annual victims has decreased over time, sexual violence appears to have increased in 2021 maintaining the largest sex difference, skewing drastically female.
-These results have implications for how Torontonians view their safety and consider how their sex impacts the possibility of crimes they fall victim to.
+By examining the data I found that in all categories of crime (assault, robbery and other) but sexual assault, the number of annual victims has decreased over time.
+Sexual violence appears to have increased in 2021 maintaining the largest sex difference, skewing drastically female.
+These results have implications for how Torontonians view their safety and consider how sex impacts their predisposition to certain crimes.
 toc: true
@@
 
-Code and data supporting this analysis is available at: LINK
+^[Code and data supporting this analysis is available at: LINK]
 
@@
 
-Prior to the 2006 Canadian federal election, Stephen Harper, who would eventually be elected Prime Minister until 2016, led a conservative campaign based on crime prevention (cite SH book).
+Prior to the 2006 Canadian federal election, Stephen Harper, who would eventually be elected Prime Minister until 2016, led a conservative campaign based on crime prevention [@book].
 The conservative party asserted that Canada was facing a period of increased crime and assured protection to Canadians by way of tougher sentencing/ parole, additional police funding, and investments in victim services, among a myriad of other things [@book].
@@
 In this paper, I utilize Open Data Toronto [@opendatatoronto] to assess the number of victims of crimes against the person in Toronto across 2014 - 2021.
-This is to investigate how the number of victims of crime in Toronto has changed over time, and if it has increased  over time, which categories of crimes against the person are most prevalent and if sex differences are apparent across different forms of crime.
+This is to investigate how the number of victims of crime in Toronto has changed over time, if it is increasing, which categories of crimes are most prevalent and if sex differences are apparent across different forms of crime.
 I will commence with a discussion of my data source and its features.
 I will then conduct a breakdown of the average number of victimization reports for the sub-categories "assault", "robbery", "sexual violence" and "other", across 2014-2019.
-Next, I will move towards an analysis of how the reported number of victims has changed according to the category of crime, over time.
-I will conclude my analysis with an investigation of how the number of victimization reports for each category of crime, differs according to the gender of the victim.
+Next, I will move towards an analysis of how the number of victims has changed according to the category of crime, over time.
+Following, I will conduct an investigation of how the number of victimization reports for each category of crime, differs according to the sex of the victim.
 The paper will culminate with a discussion of my findings, limitations and future research possibilities.
@@
 
-The data source utilized for the paper was taken from the City of Toronto's open data portal [@opendatatoronto].
+The data source utilized for this paper was taken from the City of Toronto's open data portal [@opendatatoronto].
 The data was collected from the Toronto Police Service Annual Statistical Report (ASR) which conducts overviews of police-related statistics.
@@
 I would have been forced to remove this data, thus negatively impacting my analysis.
-Because my analysis is focused on victims of crime over time and potential gender differences, I have decided to withdraw from using data for age group or cohort.
+Because I am focused on victims of crime over time and potential gender differences, I have withdrawn from using data on age group or cohort.
 
 \
-Included within the crime category subtype are (1) assault, (2) robbery, (2) sexual violation and (4) other.
+Included within the crime category subtype are (1) assault, (2) robbery, (3) sexual violation and (4) other.
 To further discuss these categorizations, in Canada, assault is classified as intentionally applying force to an individual indirectly or directly without consent.
@@
 I first selected the variables of interest, checked for missing values, cross-tabulated to understand the pattern of unidentified data and dropped unknown variables.
-Included in the sex variable was "U" for unknown, which unfortunately left me forced to remove it from my dataset.
-This is a limitation of my analysis that I will further examine in my discussion.
+Included in the sex variable was "U" for unknown, which I was unfortunatelyforced to remove it from my dataset.
+This is a limitation of my analysis that will be further examined in my discussion.
 
 My investigation began with an overview of the average number of people victimized by crime in Toronto across 2014-2021 based on the sub-category of crime.
-My intent with this is to help visualize the numerical average of Torontonians who were the victims of each sub-category of crime across 2014 - 2021.
-I visualized this in a table (see table 1) to create categorical organization from the highest to lowest average victims of crime.
+I visualized this data to create categorical organization from the highest to lowest average victims of crime, see table 1.
 
@@
              "simple", 
-             caption = "Average Number of victims of crimes in Toronto across 2014                         -2021, grouped by category of crime", 
+             caption = "Average number of victims of crimes in Toronto across 2014                         -2021, grouped by category of crime", 
              col.names=c("Category of Crime", "Average Number of Victims"), 
@@
 ```
+From this table, it is apparent that assault has the highest average number of victims total spanning 2014 - 2021.
 
@@
              "simple", 
-             caption = "The first 6 rows of the data frame used to represent the                           gender of the number of victims according to category of                           crime in Toronto Across 2014 - 2021", col.names=c("Sub                             -Category of Crime", "Sex", "Year", "Number of Victims"),
+             caption = "The first 6 rows of the data frame used to represent the                           sex of the number of victims according to category of crime in Toronto Across 2014 - 2021", col.names=c("Sub-Category of Crime", "Sex", "Year", "Number of Victims"),
              align=c('l', 'c', 'c'))
@@
 As per my primary objective for this paper, I wish to understand how the number of victims of crime in Toronto has changed over time and to analyze if victimization numbers are increasing over time.
-Because of this, I created a line graph that showcases the number of victims according to each subcategory of crime over 2014 -2019 (See Figure 2).
+Because of this, I created a line graph that showcases the number of victims according to each subcategory of crime over 2014 - 2019 (See Figure 2).
 To do this, I called on the data frame visualized above (See Table 2).
 This visualization allows for comparison across the different subcategories of crimes while simultaneously presenting a rich depiction of how the number of victims has changed over time, the year they peaked and at their lowest.
-Using the data previewed in Table 1, I was able to discriminate the number of total victims for the peak years of each sub-category of crime.
+Using the data previewed in Table 2, I was able to discriminate the number of total victims for the peak years of each sub-category of crime.
 
@@
 
-By drawing your attention to Figure 1, it is apparent from 2014-2021, the most prevalent crime that Torontonians were victims of was "Assault".
+By drawing your attention to Figure 1, it is apparent from 2014 - 2021, the most prevalent crime that Torontonians were victims of was "Assault".
 From 2014 - 2018, the number of annual victims gradually increased until it peaked in 2019, with 18819 total victims.
@@
 
-"Robbery" had the third most victims of crime from 2014-2021.
+"Robbery" had the third most victims of crime from 2014 - 2021.
 The peak number of victims for this variable is not very distinct, but saw its highest in 2017 with 3686 victims that year.
@@
 
-"Sexual Violence" had the least number of victims of crime from 2014-2021.
+"Sexual Violence" had the least number of victims of crime from 2014 - 2021.
 However, contrary to the others, sexual violence saw its peak in 2021 with 2568 victims.
@@
 Each plot assesses the number of victims across 2014 - 2021 and the number of victims that were male (M) and female (F), respectively.
-Using the data set from Table 2, which was previewed above, I determined the number of victims that were male and female for the peak years of each sub-category of crime and calculated percentage differences.
+Using the data set from Table 3, which was previewed above, I determined the number of victims that were male and female for the peak years of each sub-category of crime and calculated percentage differences.
 
@@
 Lastly, "Sexual Violence", on the bottom right, shows the most apparent sex difference from 2014-2021, where more female Torontonians were victims of sexual violence than men.
-This sex difference was the most defined across all the sub-categories, with 2244 of the total 2568 victims in its peak year, 2021 being female, and 334 being male.
+This sex difference was the most defined across all the sub-categories, with 2244 of the total 2568 victims in its peak year, 2021, being female, and 334 being male.
 This means a difference of 147.9%.
@@
 
-In this paper, I have utilized The City of Torontos open data portal, supplied by the Toronto Police Service Annual Statistical Report (ASR) to assess the number of victims of crimes against the person in Toronto across 2014-2021.
-The primary objective of this paper was to assess if the number of Torontonians victimized by crime has changed and if it has increased over time, which categories of crimes against the person are most prevalent and if sex differences are apparent across the categories of crime.
-By utilizing the Coding language R (cite) I was able to import, clean and asses the raw data made available by the City.
+In this paper, I have utilized The City of Torontos open data portal [@opendatatoronto], supplied by the Toronto Police Service Annual Statistical Report (ASR) to assess the number of victims of crimes against the person in Toronto across 2014 - 2021.
+The primary objective of this paper was to assess if the number of Torontonians victimized by crime has changed, if it has increased over time, which categories of crimes against the person are most prevalent and if sex differences are apparent.
+By utilizing the Coding language R [@R] I was able to import, clean and asses the raw data made available by the City.
 In doing so I discovered that for the sub-category of crimes, assault, robbery, and other, the number of Torontonian victims peaked in the years 2019, 2017, and 2015 respectively and decreased until the end of the data collection date (2021).
-The most prevalent crime that Torontonians were victims of was assault.
+The most prevalent crime that Torontonians were victims of was assault, maintaining the highest cumulative average across 2014-2021.
 Contrarily, sexual violence saw its peak in 2021, meaning it has not decreased over time.
@@
 The latter is an extremely apparent phenomenon for women who attempt to report sexual violence.
-While this data represents the number of individuals who have fallen victim to crimes against the person across 2014-2021, it cannot account for unfiled and dismissed reports.
+While this data represents the number of individuals who have fallen victim to crimes against the person across 2014 - 2021, it cannot account for unfilled and dismissed reports.
 Meaning that the number of victims could potentially be far higher than portrayed in the analysis.
 Regarding the data set itself, removing missing and unknown data will impact the analysis in various regards.
-More specifically, regarding the variable sex variable including "male", "female" and "unknown", I was forced to remove "unknown" as further insights were not provided on this denomination.
+More specifically, concerning the variable sex including the categories "male", "female" and "unknown", I was forced to remove "unknown" as further insights were not provided on this denomination.
 Removing this data could have prevented insightful analysis of non-binary and Agender individuals.

### 2023-02-03 outputs/paper/Paper 1.qmd
Toronto Victimization: Fifth commit, includes all updated sub-folders and final deliverable

@@
 
-^[Code and data supporting this analysis is available at: LINK]
+^[Code and data supporting this analysis is available at: https://github.com/SofiaSellitto/Toronto-Victimization-Rates]
 
@@
              "simple", 
-             caption = "Average number of victims of crimes in Toronto across 2014                         -2021, grouped by category of crime", 
+             caption = "Average number of victims of crimes in Toronto across 2014-2021, grouped by category of crime", 
              col.names=c("Category of Crime", "Average Number of Victims"), 
@@
              "simple", 
-             caption = "The first 6 rows of the data frame used to represent the                           sex of the number of victims according to category of crime in Toronto Across 2014 - 2021", col.names=c("Sub-Category of Crime", "Sex", "Year", "Number of Victims"),
-             align=c('l', 'c', 'c'))
+             caption = "The first 6 rows of the data frame used to represent thesex of the number of victims according to category of crime in Toronto Across 2014 - 2021", col.names=c("Sub-Category of Crime", "Sex", "Year", "Number of Victims"),
+             align=c('l', 'c', 'c', 'c'))
 

### 2023-02-03 inputs/data/00-download_data.R

@@ -1,4 +1,7 @@
 
+## Importing and loading raw data from The City of Toronto ##
+
+
 library(opendatatoronto)
 library(dplyr)
 

### 2023-02-03 outputs/paper/Paper 1.qmd
Toronto Victimization: Seventh commit, including all updated inputs, outputs and scripts

@@
 By examining the data I found that in all categories of crime (assault, robbery and other) but sexual assault, the number of annual victims has decreased over time.
-Sexual violence appears to have increased in 2021 maintaining the largest sex difference, skewing drastically female.
+Sexual violation appears to have increased in 2021 maintaining the largest sex difference, skewing drastically female.
 These results have implications for how Torontonians view their safety and consider how sex impacts their predisposition to certain crimes.
@@
 I will commence with a discussion of my data source and its features.
-I will then conduct a breakdown of the average number of victimization reports for the sub-categories "assault", "robbery", "sexual violence" and "other", across 2014-2019.
+I will then conduct a breakdown of the average number of victimization reports for the sub-categories "assault", "robbery", "sexual violation" and "other", across 2014-2019.
 Next, I will move towards an analysis of how the number of victims has changed according to the category of crime, over time.
@@
 In Canada, robbery is defined as stealing for extortion of the article stolen and can include the use of violent threats, personal violence or with a weapon.
-In Ontario, sexual violence describes any physical or psychological violence conducted or targeted in a sexual means.
+In Ontario, sexual violation describes any physical or psychological violence conducted or targeted in a sexual means.
 Unfortunately, no information was provided by the City of Toronto nor the Toronto Police Service Annual Statistical Report (ASR) on the nature of the "other" category.
@@
 
-See **Figure 1**
-
 ```{r}
@@
 
-"Sexual Violence" had the least number of victims of crime from 2014 - 2021.
-However, contrary to the others, sexual violence saw its peak in 2021 with 2568 victims.
+"sexual violation" had the least number of victims of crime from 2014 - 2021.
+However, contrary to the others, sexual violation saw its peak in 2021 with 2568 victims.
 From 2014 onward, the number of victims gradually increased until it peaked in 2021.
 
-In representing how sex impacts the victims of crimes, I chose to display this data using 4 bar charts based on the categories "Assault", "Other", "Robbery" and "Sexual Violence".
+In representing how sex impacts the victims of crimes, I chose to display this data using 4 bar charts based on the categories "Assault", "Other", "Robbery" and "sexual violation".
 Each plot assesses the number of victims across 2014 - 2021 and the number of victims that were male (M) and female (F), respectively.
@@
 \
-See **Figure 2**
 
@@
 ```
-
 In drawing your attention to Figure 2 contained in the "Assault", on the top left, it is apparent that from 2014 - 2021, more male Torontonians were victims of assault than women.
@@
 
-\
-Lastly, "Sexual Violence", on the bottom right, shows the most apparent sex difference from 2014-2021, where more female Torontonians were victims of sexual violence than men.
+Lastly, "sexual violation", on the bottom right, shows the most apparent sex difference from 2014-2021, where more female Torontonians were victims of sexual violation than men.
 This sex difference was the most defined across all the sub-categories, with 2244 of the total 2568 victims in its peak year, 2021, being female, and 334 being male.
@@
 The most prevalent crime that Torontonians were victims of was assault, maintaining the highest cumulative average across 2014-2021.
-Contrarily, sexual violence saw its peak in 2021, meaning it has not decreased over time.
-In assessing how sex affects the number of victims for each sub-category of crime, it was apparent that sexual violence had the largest difference between male and female victims, skewing drastically female, maintaining a difference of 147.97%.
+Contrarily, sexual violation saw its peak in 2021, meaning it has not decreased over time.
+In assessing how sex affects the number of victims for each sub-category of crime, it was apparent that sexual violation had the largest difference between male and female victims, skewing drastically female, maintaining a difference of 147.97%.
 Robbery maintains the second largest difference between male and female victims, skewing towards men with a difference of 101.35%.
@@
 Another vital point to acknowledge is the number of individuals who may have fallen victim to a crime against the person and failed to report it or those who intended to file a report but were dismissed.
-The latter is an extremely apparent phenomenon for women who attempt to report sexual violence.
+The latter is an extremely apparent phenomenon for women who attempt to report sexual violation.
 While this data represents the number of individuals who have fallen victim to crimes against the person across 2014 - 2021, it cannot account for unfilled and dismissed reports.
@@
 
-\
-
 ## References
